---
title: "Exercise_11"
output: html_document
date: "2024-04-01"
---
Before we get started, we will load all the packages needed and load the data set.
```{r}
new_pkg <- c("tidyverse", "dplyr", "mosaic", "ggplot2", "infer", "skimr", "cowplot", "broom", "naniar", "AICcmodavg", "MuMIn")
not_installed <- new_pkg[!new_pkg %in% rownames(installed.packages())]
if (length(not_installed) > 0) install.packages(not_installed) 

library("tidyverse")
library("dplyr")
library("mosaic")
library("ggplot2")
library("infer")
library("skimr")
library("cowplot")
library("broom")
library("naniar")
library("AICcmodavg")
library("MuMIn")
```

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f)


skim(d)
```
The output indicates 4 variables are categorical and 10 are numeric. They were listed in the following output. 
Column type frequency:           
  character                4     
  numeric                  10  

Challenge 1
Step 1
```{r}
d <- replace_with_na_all(data = d,
                    condition = ~.x == -999)
```

Step 2
```{r}
d_2 <- d %>%
  select(-c("litter size", "refs"))
```

Step 3
```{r}
d_3 <- d_2 %>%
  select_if(is.numeric)
```

Spaces and parentheses sometimes interfere with the codes so I rename the variables.
```{r}
names(d_3) <- c("mass", "gestation", "newborn", "weaning", "wean.mass", "AFR", "max.life", "litters.year")
```


```{r}
d_3[paste0("log.", names(d_3))] <- lapply(d_3, function(x) log(x))
```

This is the data frame with all log transformed variables
```{r}
d_log <- bind_cols(d_2, d_3)
```

Step 4
```{r}
relGest <- lm(log.gestation~log.mass, d_log, na.action=na.exclude)
relWean <- lm(log.weaning~log.mass, d_log, na.action=na.exclude)
relAFR <- lm(log.AFR~log.mass, d_log, na.action=na.exclude)
relLife <- lm(log.max.life~log.mass, d_log, na.action=na.exclude)
relNewbornMass <- lm(log.newborn~log.mass, d_log, na.action=na.exclude)
relWeaningMass <- lm(log.wean.mass~log.mass, d_log, na.action=na.exclude)
d_log <- d_log %>%
    mutate(relGest = residuals(relGest), 
           relWean = residuals(relWean), 
           relAFR = residuals(relAFR), 
           relLife = residuals(relLife), 
           relNewbornMass = residuals(relNewbornMass), 
           relWeaningMass = residuals(relWeaningMass))

```

Step 5
```{r}
p1 <- ggplot(d_log, aes(order,relLife)) +
    geom_boxplot()+
    theme(axis.text.x = element_text(size = 5,angle = 270))
p2 <- ggplot(d_log, aes(order,relNewbornMass)) +
    geom_boxplot()+
    theme(axis.text.x = element_text(size = 5,angle = 270))
p3 <- ggplot(d_log, aes(order,relWeaningMass)) +
    geom_boxplot()+
    theme(axis.text.x = element_text(size = 5,angle = 270))

plot_grid(p1,p2,p3)
```
The mammalian order with the highest residuals in terms of means:
relLife: Primates;
relNewbornMass: Cetacea;
relWeaningMass: Perissodactyla

Step 6
Here we remove NAs for models for each of the two response variables, max. life(mo) and AFR(mo)
```{r}
d_max.life <- d_log %>%
  drop_na(log.max.life, log.gestation, log.newborn, log.weaning, log.wean.mass, log.litters.year, log.mass)
```

```{r}
d_AFR <- d_log %>%
  drop_na(log.AFR, log.gestation, log.newborn, log.weaning, log.wean.mass, log.litters.year, log.mass)
```

For log.max.life:
```{r}
m0 <- lm(data = d_max.life, log.max.life ~ log.gestation + log.newborn + log.weaning + log.wean.mass + log.litters.year + log.mass)
summary(m0)
```
```{r}
drop1(m0, test = "F")
```
```{r}
m1 <- update(m0, . ~ . - log.wean.mass)
summary(m1)
```


```{r}
drop1(m1, test = "F")
```

```{r}
m2 <- update(m1, . ~ . - log.newborn)
summary(m2)
```
```{r}
drop1(m2, test = "F")
```

```{r}
aictab(list(m0, m1, m2), c("m0", "m1", "m2"))
```
Overall, m2 (dropped log.wean.mass and log.newborn) is the best model, all three models have a Delta_AICc lower than 4. log.gestation, log.weaning, log.litters.year and log.mass appear in all of this set of “top” models.

```{r}
m <- lm(data = d_max.life, log.max.life ~ log.gestation + log.newborn + log.weaning + log.wean.mass + log.litters.year + log.mass, na.action = "na.fail")
mods <- dredge(m)
```
```{r}
(mods.avg <- summary(model.avg(mods, subset = delta <= 4, fit = TRUE)))

```

```{r}
confint(mods.avg)
```

```{r}
plot(mods.avg, full = TRUE)
```
For AFR:
```{r}
m0_AFR <- lm(data = d_AFR, log.AFR ~ log.gestation + log.newborn + log.weaning + log.wean.mass + log.litters.year + log.mass)
summary(m0_AFR)
```

```{r}
drop1(m0_AFR, test = "F")
```

```{r}
m1_AFR <- update(m0_AFR, . ~ . - log.wean.mass)
summary(m1_AFR)

```

```{r}
drop1(m1_AFR, test = "F")
```

```{r}
m2_AFR <- update(m1_AFR, . ~ . - log.newborn)
summary(m2_AFR)
```

```{r}
drop1(m2_AFR, test = "F")
```

```{r}
aictab(list(m0_AFR, m1_AFR, m2_AFR), c("m0_AFR", "m1_AFR", "m2_AFR"))
```
With same variables dropped, m2_AFR is the best model for AFR, all three models have a Delta_AICc lower than 4. log.gestation, log.weaning, log.litters.year and log.mass (the same four variables) appear in all of this set of “top” models.
```{r}
m_AFR <- lm(data = d_AFR, log.AFR ~ log.gestation + log.newborn + log.weaning + log.wean.mass + log.litters.year + log.mass, na.action = "na.fail")
mods_AFR <- dredge(m_AFR)
```

```{r}
(mods.avg_AFR <- summary(model.avg(mods_AFR, subset = delta <= 4, fit = TRUE)))

```

```{r}
confint(mods.avg_AFR)
```

```{r}
plot(mods.avg_AFR, full = TRUE)
```